---
title: "Persona assignment from open data"
author: "Louison Thépaut (IVM, VU Amsterdam)"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<!--
============================================================================
ADDRESSES IN, PERSONAS OUT.

This file is self-contained: it does not source any other script. It reads
  1. the trained random forest shipped with the repository (models/persona_rf.rds),
  2. open data: CBS "Kerncijfers per postcode" (three Excel files for the
     demographics year, two for the income year) and the BAG geopackage,
  3. your address list (CSV or XLSX: postcode, house number, BAG ID),
and writes one row per address with the features used and the predicted persona.

The open data is read ONCE and cached as CSV files in cache_dir (the BAG
geopackage is the slow step, a few minutes). Every later run just reads the
cache. Set rebuild_cache <- TRUE when you download a new CBS release.

What to edit: the CONFIG chunk. Where to download the data: README.md.
Run it with Knit (RStudio) or chunk by chunk; both work.

How an address becomes a persona
--------------------------------
BAG: matched by nummeraanduiding ID, then verblijfsobject ID, then postcode +
house number. Gives floor area and construction year.
CBS: per variable, the PC6 value if published, else PC5, else PC4. Missing-value
codes (-99997 ...) are never treated as numbers.
Income: z-scored on the CBS postcode mean/SD of income_year. The survey side of
the model was z-scored on its own mean/SD, so a split in the forest means
"this many SDs above the population mean" on both sides.
Anything still missing gets a population fall-back and is flagged per row.
The forest predicts four classes; predicted Erratic (Class3) is folded into
Comfort (Class4) for deployment. Both columns are in the output.
============================================================================
-->

```{r config}
## ---- CONFIG: the only chunk you need to edit -----------------------------

# The PERSONA_* environment variables override the defaults below; the Docker
# image uses them. In RStudio, just edit the defaults.
env <- function(name, default) { v <- Sys.getenv(name); if (nzchar(v)) v else default }

# Folder with the open-data downloads (sub-folders are fine, keep file names)
data_dir    <- env("PERSONA_DATA_DIR", "~/persona-data")
cbs_year    <- as.integer(env("PERSONA_CBS_YEAR", 2025))     # CBS demographics: pc6/pc5/pc4_<year>_*.xlsx
income_year <- as.integer(env("PERSONA_INCOME_YEAR", 2023))  # CBS household income: pc5/pc4_<year>_vol.xlsx (normally cbs_year - 2)
bag_gpkg    <- file.path(data_dir, "bag-light.gpkg")
cache_dir   <- file.path(data_dir, "cache")   # CSV cache written on the first run
rebuild_cache <- as.logical(env("PERSONA_REBUILD_CACHE", FALSE))  # TRUE = read the raw files again

# Your addresses: a CSV/XLSX path (relative paths are taken from the repository
# folder), or a data.frame typed straight in, e.g.
# addresses <- data.frame(postcode = c("1011AB", "3011AD"), huisnummer = c("12", "5A"),
#                         bag_id = c("0363200000123456", "0599200000123456"))
addresses       <- env("PERSONA_ADDRESSES", "examples/addresses_example.csv")
addresses_sheet <- NULL      # XLSX only: sheet name or number, NULL = first sheet
addresses_skip  <- 0         # rows above the header row
address_col_map <- list()    # force a column if auto-detection picks the wrong one,
                             # e.g. list(bag_id = "Woningcode", postcode = "Postcode 6")

# Output (relative paths are taken from the repository folder)
output_dir       <- env("PERSONA_OUTPUT_DIR", "output")
path_output_csv  <- file.path(output_dir, "persona_predictions.csv")    # everything: features, diagnostics, personas
path_output_xlsx <- file.path(output_dir, "persona_predictions.xlsx")   # same table as a workbook; NULL to skip

# The repository folder is found automatically (see the locate chunk). Only
# set it here if that fails, e.g. repo_dir <- "~/persona-segmentation".
repo_dir <- env("PERSONA_REPO_DIR", NULL)
## ---------------------------------------------------------------------------
```

```{r libraries}
required_packages <- c("tidyverse", "data.table", "janitor", "readxl", "writexl", "ranger")
to_install <- setdiff(required_packages, rownames(installed.packages()))
if (length(to_install) > 0) install.packages(to_install)
# "sf" is only needed the first time, to read the BAG geopackage; installed then.

library(tidyverse)
library(data.table)
library(janitor)
library(readxl)
library(writexl)
library(ranger)
```

```{r locate}
# The repository folder: where this Rmd, models/persona_rf.rds and examples/ live.
if (is.null(repo_dir)) {
  repo_dir <- tryCatch(dirname(normalizePath(knitr::current_input(dir = TRUE))), error = function(e) NULL)
  if (is.null(repo_dir) && requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable())
    repo_dir <- tryCatch(dirname(rstudioapi::getSourceEditorContext()$path), error = function(e) NULL)
  if (is.null(repo_dir) || repo_dir == "") repo_dir <- getwd()
}
repo_dir   <- path.expand(repo_dir)
model_path <- file.path(repo_dir, "models", "persona_rf.rds")
if (!file.exists(model_path)) {
  stop("Cannot find models/persona_rf.rds under ", repo_dir,
       ". Set repo_dir in the config chunk to the folder you cloned from GitHub.", call. = FALSE)
}
from_repo <- function(p) if (is.null(p) || grepl("^(/|~|[A-Za-z]:)", p)) path.expand(p) else file.path(repo_dir, p)
cat("Repository folder:", repo_dir, "\n")
```

# Functions

Everything the pipeline needs, in one place. You should not have to touch this
chunk unless CBS renames a column (then edit the pattern in `CBS_VARS`).

```{r functions}
# ---- constants ---------------------------------------------------------------

# Features in the order the forest expects them (also stored in the model bundle).
MODEL_FEATURES <- c("female", "age", "members", "year_built", "rental", "urban",
                    "nettohh_z", "square_meters")

# LISS labels of the construction-period variable. The training template checks
# that the survey data really uses these; the field side is labelled from the
# levels stored in the model bundle, so the two can never drift apart.
# "\u00f3" spells the accented o so the file works in any locale
YEAR_BUILT_LEVELS <- c("V\u00f3\u00f3r 1940", "Tussen 1940 en 1970",
                       "Tussen 1971 en 2000", "In 2001 of later")
YEAR_BUILT_BREAKS <- c(1940, 1971, 2001)   # bouwjaar < 1940 | 1940-1970 | 1971-2000 | >= 2001

CLASS_LABELS <- c(Class1 = "Financially", Class2 = "Policy",
                  Class3 = "Erratic",     Class4 = "Comfort")

# Deployment rule: Erratic choosers receive the Comfort letter. This is applied
# AFTER the forest has been trained and evaluated as a four-class model.
FOLD_RULE <- c(Class3 = "Class4")

CBS_SENTINEL_LIMIT <- -9000              # values below this are CBS missing codes
BAG_AREA_RANGE     <- c(10, 2000)        # m2; outside -> NA (BAG placeholders: 1, 999999)
BAG_YEAR_RANGE     <- c(1500, as.integer(format(Sys.Date(), "%Y")) + 2)

# CBS variables we read, matched by name on the cleaned Excel header. If CBS
# renames a column, edit the pattern here (the error message lists what it found).
#   to = what a missing-value code becomes: 0 for counts, NA for anything else
CBS_VARS <- list(
  man             = list(pattern = "^(aantal_)?man(nen)?$",                 required = TRUE,  to = 0),
  vrouw           = list(pattern = "^(aantal_)?vrouw(en)?$",                required = TRUE,  to = 0),
  age_0_15        = list(pattern = "tot_15_jaar$",                          required = TRUE,  to = 0),
  age_15_25       = list(pattern = "15_tot_25_jaar$",                       required = TRUE,  to = 0),
  age_25_45       = list(pattern = "25_tot_45_jaar$",                       required = TRUE,  to = 0),
  age_45_65       = list(pattern = "45_tot_65_jaar$",                       required = TRUE,  to = 0),
  age_65_plus     = list(pattern = "65_jaar_(en|of)_ouder",                             required = TRUE,  to = 0),
  huishoudgrootte = list(pattern = "huishoud(ens)?grootte",                 required = TRUE,  to = NA),
  huurwoning      = list(pattern = "^(percentage_|p_|aandeel_)?huurwoning(en)?$", required = TRUE, to = NA),
  koopwoning      = list(pattern = "^(percentage_|p_|aandeel_)?koopwoning(en)?$", required = FALSE, to = NA),
  woningen        = list(pattern = "^(aantal_)?woningen$",                  required = FALSE, to = 0),
  woz             = list(pattern = "woz",                                   required = FALSE, to = NA),
  inkomen         = list(pattern = "inkomen",                               required = FALSE, to = NA),
  stedelijkheid   = list(pattern = "stedelijk",                             required = FALSE, to = NA)
)

# ---- small utilities ---------------------------------------------------------

say <- function(...) cat(sprintf(...), "\n", sep = "")

`%||%` <- function(a, b) if (is.null(a)) b else a

normalise_postcode <- function(x) {
  x <- toupper(gsub("\\s+", "", as.character(x)))
  x[is.na(x) | x == ""] <- NA_character_
  x
}

is_valid_pc6 <- function(x) !is.na(x) & grepl("^[1-9][0-9]{3}[A-Z]{2}$", x)

# Turn anything (numeric, text, "1.9632e+15", "01963200000012345") into a clean
# 16-digit BAG ID or NA. Returns a character vector; attr "status" says what
# happened to each element.
normalise_bag_id <- function(x, width = 16L) {
  raw <- if (is.numeric(x)) sprintf("%.0f", x) else as.character(x)
  raw <- trimws(raw)
  raw[is.na(raw) | raw == "" | raw == "NA"] <- NA_character_

  status <- rep("ok", length(raw))
  status[is.na(raw)] <- "missing"

  # Scientific notation means the digits are already gone: refuse to guess.
  sci <- !is.na(raw) & grepl("[eE][+-]?[0-9]", raw)
  status[sci] <- "invalid (scientific notation, digits lost)"
  raw[sci] <- NA_character_

  x <- sub("\\.0+$", "", raw)          # "363200000123456.0"
  x <- gsub("[^0-9]", "", x)           # keep digits only

  short <- !is.na(x) & nchar(x) < width & nchar(x) > 0
  x[short] <- str_pad(x[short], width, side = "left", pad = "0")
  status[short] <- "padded (leading zeros restored)"

  long <- !is.na(x) & nchar(x) > width
  x[long] <- sub(paste0("^0+(?=[0-9]{", width, "}$)"), "", x[long], perl = TRUE)
  status[long & !is.na(x) & nchar(x) == width] <- "trimmed (spurious leading zeros removed)"

  bad <- !is.na(x) & nchar(x) != width
  status[bad] <- "invalid (not 16 digits)"
  x[bad] <- NA_character_
  x[!is.na(x) & x == ""] <- NA_character_

  attr(x, "status") <- status
  x
}

# BAG object type sits in digits 5-6 of the ID.
bag_id_type <- function(x) {
  t <- substr(x, 5, 6)
  dplyr::case_when(t == "20" ~ "nummeraanduiding", t == "01" ~ "verblijfsobject",
                   t == "10" ~ "pand", t == "02" ~ "ligplaats", t == "03" ~ "standplaats",
                   TRUE ~ "unknown")
}

# "12", "12A", "12-D", "10-2", "12 A 3" -> number, letter, toevoeging
parse_huisnummer <- function(x) {
  x <- trimws(as.character(x))
  m <- str_match(x, "^([0-9]+)\\s*[-/]?\\s*([A-Za-z]?)\\s*[-/]?\\s*(.*)$")
  tibble(huisnummer = suppressWarnings(as.integer(m[, 2])),
         huisletter = tolower(m[, 3]),
         toevoeging = tolower(gsub("[^A-Za-z0-9]", "", m[, 4])))
}

address_key <- function(postcode, huisnummer, huisletter = "", toevoeging = "") {
  hl <- tolower(trimws(coalesce(as.character(huisletter), "")))
  tv <- tolower(gsub("[^A-Za-z0-9]", "", coalesce(as.character(toevoeging), "")))
  key <- paste(normalise_postcode(postcode), as.integer(huisnummer), hl, tv, sep = "|")
  key[is.na(postcode) | is.na(huisnummer)] <- NA_character_
  key
}

cbs_num <- function(x, sentinel_to = NA_real_) {
  x <- as.character(x)
  # Some CBS columns (income in the "vol" files) are text with a decimal comma: "41,0"
  x <- sub("^(-?[0-9]+),([0-9]+)$", "\\1.\\2", x)
  x <- suppressWarnings(as.numeric(x))
  x[!is.na(x) & x < CBS_SENTINEL_LIMIT] <- sentinel_to
  x
}

# ---- CBS postcode files -------------------------------------------------------

# Find pc<level>_<year>_*.xlsx anywhere under data_dir. Prefers the most
# complete release (vol > v2 > v1) when several exist.
find_cbs_file <- function(data_dir, level, year) {
  pat   <- sprintf("^(.*[_-])?pc%d[_-]%d[^/]*\\.xlsx?$", level, year)   # pc6_2025_v1.xlsx, cbs_pc6_2025_v1.xlsx, ...
  files <- list.files(data_dir, pattern = pat, recursive = TRUE, full.names = TRUE,
                      ignore.case = TRUE)
  files <- files[!grepl("^~\\$", basename(files))]          # Excel lock files
  if (length(files) == 0) {
    stop(sprintf(paste(
      "No CBS file for PC%d, year %d found under '%s'.",
      "Download 'Kerncijfers per postcode' from",
      "https://www.cbs.nl/nl-nl/dossier/nederland-regionaal/geografische-data/gegevens-per-postcode",
      "unzip it anywhere inside that folder, and keep the original file name (e.g. pc%d_%d_v2.xlsx)."),
      level, year, data_dir, level, year), call. = FALSE)
  }
  rank <- ifelse(grepl("vol", basename(files), ignore.case = TRUE), 3,
          ifelse(grepl("v2",  basename(files), ignore.case = TRUE), 2, 1))
  files[order(-rank, basename(files))][1]
}

# The CBS workbooks have a title block, then a header row, then a units row,
# then the data. Row positions differ between releases, so we look for the row
# whose first cell starts with "Postcode" instead of hard-coding skip = 8.
detect_cbs_header <- function(preview) {
  first <- tolower(trimws(as.character(preview[[1]])))
  h <- which(grepl("^(postcode|pc[456]\\b)", first))[1]
  if (is.na(h)) {
    # maybe the postcode column is not the first one
    hits <- which(apply(preview, 1, function(r) any(grepl("^postcode", tolower(trimws(r))))))
    h <- hits[1]
  }
  if (is.na(h)) stop("Could not find the header row (a row starting with 'Postcode') in the CBS file.", call. = FALSE)
  h
}

# Match CBS_VARS patterns against cleaned header names. Returns a named integer
# vector of column positions (NA for optional variables that are absent).
match_cbs_columns <- function(clean_names, level) {
  idx <- sapply(names(CBS_VARS), function(v) {
    hits <- which(grepl(CBS_VARS[[v]]$pattern, clean_names, ignore.case = TRUE))
    if (v == "inkomen" && length(hits) > 1) hits <- hits[grepl("gemiddeld|gem_", clean_names[hits])]
    if (length(hits) == 1) return(hits)
    if (length(hits) == 0) return(NA_integer_)
    stop(sprintf("Several columns match '%s' in the PC%d file: %s. Tighten CBS_VARS$%s$pattern in the functions chunk.",
                 v, level, paste(clean_names[hits], collapse = ", "), v), call. = FALSE)
  })
  missing_required <- names(CBS_VARS)[sapply(CBS_VARS, `[[`, "required") & is.na(idx)]
  if (length(missing_required) > 0) {
    stop(sprintf(paste(
      "PC%d file: could not find column(s) %s.",
      "Cleaned header names in the file are:\n  %s\n",
      "Adjust the matching pattern(s) in CBS_VARS (functions chunk of this file)."),
      level, paste(missing_required, collapse = ", "), paste(clean_names, collapse = ", ")), call. = FALSE)
  }
  idx
}

# Read one CBS postcode workbook and return a clean tibble: postcode + the
# CBS_VARS columns as numbers (missing codes handled), one row per postcode.
read_cbs_sheet <- function(path, level, sheet = 1) {
  say("  reading %s ...", basename(path))
  preview <- suppressMessages(read_excel(path, sheet = sheet, col_names = FALSE,
                                         col_types = "text", n_max = 30))
  h <- detect_cbs_header(preview)
  header <- as.character(preview[h, ])
  # Fill empty header cells from the row above or below (CBS sometimes puts the
  # label of the last column, Stedelijkheid, on another header line).
  for (j in seq_along(header)) {
    if (is.na(header[j]) || trimws(header[j]) == "") {
      cand <- c(if (h > 1) preview[[j]][h - 1] else NA, preview[[j]][h + 1])
      cand <- cand[!is.na(cand) & trimws(cand) != ""]
      if (length(cand) > 0) header[j] <- cand[1]
    }
  }
  clean <- suppressWarnings(make_clean_names(ifelse(is.na(header) | header == "", paste0("col", seq_along(header)), header)))

  idx <- match_cbs_columns(clean, level)
  # Stedelijkheid: only in PC5/PC4; CBS has shipped it with an empty header.
  if (level %in% c(4, 5) && is.na(idx["stedelijkheid"])) {
    if (grepl("^col[0-9]+$", clean[length(clean)])) {
      idx["stedelijkheid"] <- length(clean)
      say("  note: no column named 'Stedelijkheid' in the PC%d file; using the unlabelled last column (%d) as urbanity.",
          level, length(clean))
    } else {
      say("  note: no urbanity (Stedelijkheid) column in the PC%d file.", level)
    }
  }

  # Second pass: read only the columns we need, everything as text.
  keep  <- c(1L, idx[!is.na(idx)])
  types <- rep("skip", length(clean))
  types[keep] <- "text"
  if (anyDuplicated(keep)) stop("Two CBS variables matched the same column; tighten the patterns in CBS_VARS.", call. = FALSE)
  dat <- suppressMessages(read_excel(path, sheet = sheet, col_names = FALSE,
                                     col_types = types, skip = h))
  # readxl returns the kept columns in file order, so name them by ascending position
  nm  <- c("postcode", names(idx)[!is.na(idx)])
  if (ncol(dat) != length(nm)) stop("Unexpected column count when re-reading ", basename(path), call. = FALSE)
  dat <- setNames(dat, nm[order(keep)])

  pc_pattern <- c(`6` = "^[1-9][0-9]{3}[A-Z]{2}$", `5` = "^[1-9][0-9]{3}[A-Z]$", `4` = "^[1-9][0-9]{3}$")[[as.character(level)]]
  dat <- dat %>%
    mutate(postcode = normalise_postcode(postcode)) %>%
    filter(!is.na(postcode), grepl(pc_pattern, postcode)) %>%
    distinct(postcode, .keep_all = TRUE)
  for (v in setdiff(names(dat), "postcode")) dat[[v]] <- cbs_num(dat[[v]], CBS_VARS[[v]]$to)
  say("  %s: %s postcodes, columns: %s", basename(path), format(nrow(dat), big.mark = ","),
      paste(setdiff(names(dat), "postcode"), collapse = ", "))
  dat
}

# Derived features for one level (PC6, PC5 or PC4).
derive_cbs_features <- function(d) {
  pop_sex <- d$man + d$vrouw
  d$female_share <- ifelse(!is.na(pop_sex) & pop_sex > 0, d$vrouw / pop_sex, NA_real_)

  pop_age <- d$age_0_15 + d$age_15_25 + d$age_25_45 + d$age_45_65 + d$age_65_plus
  d$avg_age <- ifelse(!is.na(pop_age) & pop_age > 0,
                      (d$age_0_15 * 7.5 + d$age_15_25 * 20 + d$age_25_45 * 35 +
                         d$age_45_65 * 55 + d$age_65_plus * 75) / pop_age, NA_real_)

  # CBS publishes koop/huur as percentages of the housing stock in the postcode
  # files; some custom extracts carry counts. Detect which and do the right thing.
  if ("koopwoning" %in% names(d)) {
    both <- d$huurwoning + d$koopwoning
    pct_like <- median(both[!is.na(both) & both > 0])
  } else pct_like <- NA
  if (is.finite(pct_like) && pct_like >= 70 && pct_like <= 110) {
    d$rental_share <- d$huurwoning / 100
    attr(d, "rental_basis") <- "percentage"
  } else if ("woningen" %in% names(d)) {
    d$rental_share <- ifelse(!is.na(d$woningen) & d$woningen > 0, d$huurwoning / d$woningen, NA_real_)
    attr(d, "rental_basis") <- "count / total dwellings"
  } else {
    d$rental_share <- d$huurwoning / 100
    attr(d, "rental_basis") <- "percentage (assumed)"
  }
  d$rental_share[!is.na(d$rental_share) & (d$rental_share < 0 | d$rental_share > 1)] <- NA_real_

  if ("woz" %in% names(d)) d$woz_eur <- ifelse(!is.na(d$woz) & d$woz > 0, d$woz * 1000, NA_real_)
  if ("stedelijkheid" %in% names(d)) {
    d$urban <- d$stedelijkheid
    d$urban[!is.na(d$urban) & !(d$urban %in% 1:5)] <- NA_real_
  }
  if ("inkomen" %in% names(d)) d$income_eur <- ifelse(!is.na(d$inkomen) & d$inkomen > 0, d$inkomen * 1000, NA_real_)
  d
}

# Build the three CBS tables (PC6, PC5, PC4) plus the income z-score.
# income_year may equal cbs_year if that release already carries income.
build_cbs_tables <- function(data_dir, cbs_year, income_year = cbs_year - 1) {
  say("CBS demographics, year %d", cbs_year)
  raw <- list(
    pc6 = read_cbs_sheet(find_cbs_file(data_dir, 6, cbs_year), 6),
    pc5 = read_cbs_sheet(find_cbs_file(data_dir, 5, cbs_year), 5),
    pc4 = read_cbs_sheet(find_cbs_file(data_dir, 4, cbs_year), 4))
  tabs <- lapply(raw, derive_cbs_features)
  say("  rental share basis: %s", attr(tabs$pc6, "rental_basis"))

  # Income: PC5 preferred, PC4 fallback. Only the 'vol' release carries it.
  say("CBS household income, year %d", income_year)
  inc <- list()
  for (lv in c(5, 4)) {
    f <- find_cbs_file(data_dir, lv, income_year)
    t <- if (income_year == cbs_year) raw[[paste0("pc", lv)]] else read_cbs_sheet(f, lv)
    if (!"inkomen" %in% names(t)) {
      stop(sprintf(paste(
        "The PC%d file for %d (%s) has no income column.",
        "Household income is only published in the 'vol' release, normally two years behind the demographics.",
        "Set income_year to a year for which pc5_<year>_vol.xlsx and pc4_<year>_vol.xlsx exist."),
        lv, income_year, basename(f)), call. = FALSE)
    }
    inc[[paste0("pc", lv)]] <- derive_cbs_features(t)[, c("postcode", "income_eur")]
  }
  vals <- c(inc$pc5$income_eur, inc$pc4$income_eur)
  vals <- vals[!is.na(vals)]
  moments <- c(mean = mean(vals), sd = sd(vals), n = length(vals))
  say("  income moments (PC5+PC4 pooled): mean = %.0f EUR, SD = %.0f EUR, n = %d postcodes",
      moments["mean"], moments["sd"], moments["n"])
  z <- function(e) (e - moments["mean"]) / moments["sd"]
  inc$pc5$income_z <- z(inc$pc5$income_eur)
  inc$pc4$income_z <- z(inc$pc4$income_eur)

  keep <- c("postcode", "female_share", "avg_age", "huishoudgrootte", "rental_share", "woz_eur", "urban")
  out <- list(
    pc6 = tabs$pc6[, intersect(keep, names(tabs$pc6))],
    pc5 = tabs$pc5[, intersect(keep, names(tabs$pc5))] %>% left_join(inc$pc5, by = "postcode"),
    pc4 = tabs$pc4[, intersect(keep, names(tabs$pc4))] %>% left_join(inc$pc4, by = "postcode"))
  attr(out, "income_moments") <- moments
  attr(out, "sources") <- c(pc6 = find_cbs_file(data_dir, 6, cbs_year), pc5 = find_cbs_file(data_dir, 5, cbs_year),
                            pc4 = find_cbs_file(data_dir, 4, cbs_year),
                            income_pc5 = find_cbs_file(data_dir, 5, income_year), income_pc4 = find_cbs_file(data_dir, 4, income_year))
  out
}

# ---- BAG (building registry) ---------------------------------------------------

# Extract the verblijfsobject layer of bag-light.gpkg (PDOK) without geometry.
# Returns one row per verblijfsobject with clean 16-digit IDs.
build_bag_table <- function(gpkg, layer = "verblijfsobject") {
  if (!requireNamespace("sf", quietly = TRUE)) stop("Package 'sf' is needed to read the BAG geopackage: install.packages('sf')", call. = FALSE)
  if (!file.exists(gpkg)) {
    stop(sprintf(paste(
      "BAG geopackage not found: %s",
      "Download bag-light.gpkg from the PDOK BAG Atom feed (https://service.pdok.nl/lv/bag/atom/bag.xml,",
      "direct link https://service.pdok.nl/lv/bag/atom/downloads/bag-light.gpkg, several GB) and set bag_gpkg to its path."),
      gpkg), call. = FALSE)
  }
  say("BAG: reading layer '%s' from %s (this is the slow step, a few minutes)", layer, basename(gpkg))
  probe <- suppressWarnings(sf::st_read(gpkg, query = sprintf('SELECT * FROM "%s" LIMIT 1', layer), quiet = TRUE))
  cols  <- setdiff(names(probe), attr(probe, "sf_column"))

  need <- c(identificatie = "verblijfsobject_id",
            nummeraanduiding_hoofdadres_identificatie = "bag_id",
            oppervlakte = "oppervlakte_m2",
            bouwjaar = "bouwjaar")
  opt  <- c(postcode = "postcode", huisnummer = "huisnummer", huisletter = "huisletter",
            huisnummertoevoeging = "toevoeging", toevoeging = "toevoeging",
            openbare_ruimte_naam = "straat", woonplaats_naam = "woonplaats",
            pand_identificatie = "pand_id", gebruiksdoel = "gebruiksdoel")
  miss <- setdiff(names(need), cols)
  if (length(miss) > 0) {
    stop(sprintf("Layer '%s' lacks column(s) %s. Columns present: %s. Edit 'need' in build_bag_table().",
                 layer, paste(miss, collapse = ", "), paste(cols, collapse = ", ")), call. = FALSE)
  }
  sel <- c(need, opt[names(opt) %in% cols])
  sel <- sel[!duplicated(unname(sel))]
  q   <- sprintf('SELECT %s FROM "%s"',
                 paste(sprintf('"%s" AS %s', names(sel), unname(sel)), collapse = ", "), layer)
  bag <- suppressWarnings(sf::st_read(gpkg, query = q, quiet = TRUE))
  if (inherits(bag, "sf")) bag <- sf::st_drop_geometry(bag)
  bag <- as_tibble(bag)

  bag$bag_id             <- as.character(normalise_bag_id(bag$bag_id))
  bag$verblijfsobject_id <- as.character(normalise_bag_id(bag$verblijfsobject_id))
  bag$oppervlakte_m2 <- suppressWarnings(as.numeric(bag$oppervlakte_m2))
  bag$oppervlakte_m2[!is.na(bag$oppervlakte_m2) &
                       (bag$oppervlakte_m2 < BAG_AREA_RANGE[1] | bag$oppervlakte_m2 > BAG_AREA_RANGE[2])] <- NA
  bag$bouwjaar <- suppressWarnings(as.integer(bag$bouwjaar))
  bag$bouwjaar[!is.na(bag$bouwjaar) & (bag$bouwjaar < BAG_YEAR_RANGE[1] | bag$bouwjaar > BAG_YEAR_RANGE[2])] <- NA
  if (all(c("postcode", "huisnummer") %in% names(bag))) {
    bag$address_key <- address_key(bag$postcode, bag$huisnummer,
                                   bag$huisletter %||% "", bag$toevoeging %||% "")
  }
  say("  %s verblijfsobjecten; %.1f%% with usable floor area, %.1f%% with usable construction year",
      format(nrow(bag), big.mark = ","), 100 * mean(!is.na(bag$oppervlakte_m2)), 100 * mean(!is.na(bag$bouwjaar)))
  bag
}

# ---- cache: write once, read every time --------------------------------------

cache_paths <- function(cache_dir, cbs_year) {
  list(pc6 = file.path(cache_dir, sprintf("cbs_pc6_%d.csv", cbs_year)),
       pc5 = file.path(cache_dir, sprintf("cbs_pc5_%d.csv", cbs_year)),
       pc4 = file.path(cache_dir, sprintf("cbs_pc4_%d.csv", cbs_year)),
       bag = file.path(cache_dir, "bag_verblijfsobjecten.csv"),
       meta = file.path(cache_dir, sprintf("open_data_meta_%d.csv", cbs_year)))
}

write_open_data_cache <- function(open_data, cache_dir, cbs_year) {
  dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)
  p <- cache_paths(cache_dir, cbs_year)
  for (lv in c("pc6", "pc5", "pc4")) fwrite(open_data$cbs[[lv]], p[[lv]])
  fwrite(open_data$bag, p$bag)
  fwrite(open_data$meta, p$meta)
  say("Cache written to %s", cache_dir)
  invisible(p)
}

read_open_data_cache <- function(cache_dir, cbs_year) {
  p <- cache_paths(cache_dir, cbs_year)
  rd <- function(f, char_cols) {
    hdr <- names(fread(f, nrows = 0))
    cc  <- intersect(char_cols, hdr)
    as_tibble(fread(f, colClasses = if (length(cc)) list(character = cc) else NULL, na.strings = c("", "NA")))
  }
  cbs <- list(pc6 = rd(p$pc6, "postcode"), pc5 = rd(p$pc5, "postcode"), pc4 = rd(p$pc4, "postcode"))
  bag <- rd(p$bag, c("bag_id", "verblijfsobject_id", "pand_id", "postcode", "huisletter",
                     "toevoeging", "straat", "woonplaats", "gebruiksdoel", "address_key"))
  meta <- rd(p$meta, c("key", "value"))
  list(cbs = cbs, bag = bag, meta = meta)
}

open_data_cache_exists <- function(cache_dir, cbs_year) all(file.exists(unlist(cache_paths(cache_dir, cbs_year))))

meta_value <- function(open_data, key) {
  v <- open_data$meta$value[open_data$meta$key == key]
  if (length(v) == 0) NA else v[1]
}

# One entry point for everything: returns the cache, building it if needed.
get_open_data <- function(data_dir, cbs_year, income_year = cbs_year - 1,
                          cache_dir = file.path(data_dir, "cache"),
                          bag_gpkg = file.path(data_dir, "bag-light.gpkg"), rebuild = FALSE) {
  if (!rebuild && open_data_cache_exists(cache_dir, cbs_year)) {
    say("Open-data cache found in %s (CBS %d). Reading it; set rebuild = TRUE to rebuild.", cache_dir, cbs_year)
    return(read_open_data_cache(cache_dir, cbs_year))
  }
  say("Building the open-data cache (one-off).")
  cbs <- build_cbs_tables(data_dir, cbs_year, income_year)
  bag_path <- cache_paths(cache_dir, cbs_year)$bag
  bag <- if (!rebuild && file.exists(bag_path)) {
    say("BAG cache %s already exists; reusing it.", basename(bag_path))
    read_open_data_cache_bag(bag_path)
  } else build_bag_table(bag_gpkg)

  # Population-level fall-backs used when an address has no data at any level.
  yb_idx <- findInterval(bag$bouwjaar, YEAR_BUILT_BREAKS) + 1
  yb_mode <- names(which.max(table(yb_idx)))
  mom <- attr(cbs, "income_moments")
  meta <- tibble(key = c("cbs_year", "income_year", "built_on", "income_mean_eur", "income_sd_eur", "income_n_postcodes",
                         "default_female_share", "default_avg_age", "default_huishoudgrootte", "default_rental_share",
                         "default_urban", "default_income_z", "default_square_meters", "default_year_built_index",
                         paste0("source_", names(attr(cbs, "sources")))),
                 value = as.character(c(cbs_year, income_year, format(Sys.time(), "%Y-%m-%d %H:%M"),
                                        mom["mean"], mom["sd"], mom["n"],
                                        median(cbs$pc4$female_share, na.rm = TRUE), median(cbs$pc4$avg_age, na.rm = TRUE),
                                        median(cbs$pc4$huishoudgrootte, na.rm = TRUE), median(cbs$pc4$rental_share, na.rm = TRUE),
                                        median(cbs$pc4$urban, na.rm = TRUE), 0,
                                        median(bag$oppervlakte_m2, na.rm = TRUE), yb_mode,
                                        attr(cbs, "sources"))))
  open_data <- list(cbs = cbs[c("pc6", "pc5", "pc4")], bag = bag, meta = meta)
  write_open_data_cache(open_data, cache_dir, cbs_year)
  open_data
}

read_open_data_cache_bag <- function(bag_path) {
  hdr <- names(fread(bag_path, nrows = 0))
  cc  <- intersect(c("bag_id", "verblijfsobject_id", "pand_id", "postcode", "huisletter", "toevoeging",
                     "straat", "woonplaats", "gebruiksdoel", "address_key"), hdr)
  as_tibble(fread(bag_path, colClasses = list(character = cc), na.strings = c("", "NA")))
}

print_open_data_summary <- function(open_data) {
  cov <- function(t, v) if (v %in% names(t)) sprintf("%s %.0f%%", v, 100 * mean(!is.na(t[[v]]))) else NULL
  say("\nOpen data (CBS %s, income %s, built %s)", meta_value(open_data, "cbs_year"),
      meta_value(open_data, "income_year"), meta_value(open_data, "built_on"))
  for (lv in c("pc6", "pc5", "pc4")) {
    t <- open_data$cbs[[lv]]
    say("  %s: %s postcodes | coverage: %s", toupper(lv), format(nrow(t), big.mark = ","),
        paste(unlist(lapply(c("female_share", "avg_age", "huishoudgrootte", "rental_share", "income_z", "urban"),
                            function(v) cov(t, v))), collapse = ", "))
  }
  say("  BAG: %s verblijfsobjecten, address matching %s",
      format(nrow(open_data$bag), big.mark = ","),
      if ("address_key" %in% names(open_data$bag)) "available" else "not available (no address columns in the layer)")
  say("  income moments: mean %s EUR, SD %s EUR",
      format(round(as.numeric(meta_value(open_data, "income_mean_eur"))), big.mark = ","),
      format(round(as.numeric(meta_value(open_data, "income_sd_eur"))), big.mark = ","))
}

# ---- address lists ----------------------------------------------------------

# Read a CSV or Excel address list with every column as text (so BAG IDs and
# postcodes are never mangled). Returns the sheet exactly as it is.
read_address_file <- function(path, sheet = NULL, skip = 0) {
  if (!file.exists(path)) stop("Address file not found: ", path, call. = FALSE)
  ext <- tolower(tools::file_ext(path))
  if (ext %in% c("xlsx", "xls")) {
    suppressMessages(read_excel(path, sheet = sheet %||% 1, skip = skip, col_types = "text"))
  } else if (ext %in% c("csv", "txt")) {
    suppressMessages(read_delim(path, delim = NULL, skip = skip, col_types = cols(.default = col_character()), show_col_types = FALSE))
  } else stop("Address file must be .csv, .xlsx or .xls: ", path, call. = FALSE)
}

# Find the columns we need in whatever the batch calls them. col_map can force
# any of: bag_id, postcode, huisnummer, huisletter, toevoeging.
standardise_addresses <- function(df, col_map = list()) {
  find_col <- function(key, pattern) {
    if (!is.null(col_map[[key]])) {
      if (!col_map[[key]] %in% names(df)) stop(sprintf("Column '%s' (col_map$%s) not in the address file.", col_map[[key]], key), call. = FALSE)
      return(col_map[[key]])
    }
    hits <- names(df)[grepl(pattern, names(df), ignore.case = TRUE)]
    if (length(hits) == 0) return(NULL)
    hits[1]
  }
  c_bag <- find_col("bag_id", "bag|nummeraanduiding|adresseerbaar")
  c_pc  <- find_col("postcode", "postcode|^pc6?$")
  c_hn  <- find_col("huisnummer", "huisnummer|house|^nr\\.?$|^number$")
  c_hl  <- find_col("huisletter", "huisletter")
  c_tv  <- find_col("toevoeging", "toevoeging")
  if (is.null(c_pc)) stop("No postcode column found in the address file. Columns: ", paste(names(df), collapse = ", "), call. = FALSE)
  if (is.null(c_bag)) say("note: no BAG ID column found; addresses will be matched to the BAG on postcode + house number only.")
  say("Address columns: bag_id = %s, postcode = %s, huisnummer = %s%s%s",
      c_bag %||% "(none)", c_pc, c_hn %||% "(none)",
      if (!is.null(c_hl)) paste0(", huisletter = ", c_hl) else "",
      if (!is.null(c_tv)) paste0(", toevoeging = ", c_tv) else "")

  bag_raw <- if (!is.null(c_bag)) as.character(df[[c_bag]]) else rep(NA_character_, nrow(df))
  bag_id  <- normalise_bag_id(bag_raw)
  bag_status <- attr(bag_id, "status")    # read the attribute before tibble() strips it
  hn      <- if (!is.null(c_hn)) parse_huisnummer(df[[c_hn]]) else tibble(huisnummer = NA_integer_, huisletter = "", toevoeging = "")[rep(1, nrow(df)), ]
  if (!is.null(c_hl)) hn$huisletter <- ifelse(is.na(df[[c_hl]]) | df[[c_hl]] == "", hn$huisletter, tolower(trimws(df[[c_hl]])))
  if (!is.null(c_tv)) hn$toevoeging <- ifelse(is.na(df[[c_tv]]) | df[[c_tv]] == "", hn$toevoeging, tolower(gsub("[^A-Za-z0-9]", "", df[[c_tv]])))

  work <- tibble(
    .row_id       = seq_len(nrow(df)),
    bag_id_raw    = bag_raw,
    bag_id        = as.character(bag_id),
    bag_id_status = bag_status,
    bag_id_kind   = bag_id_type(as.character(bag_id)),
    postcode_raw  = as.character(df[[c_pc]]),
    postcode      = normalise_postcode(df[[c_pc]]),
    huisnummer_raw = if (!is.null(c_hn)) as.character(df[[c_hn]]) else NA_character_,
    huisnummer    = hn$huisnummer,
    huisletter    = hn$huisletter,
    toevoeging    = hn$toevoeging)
  work$postcode_valid <- is_valid_pc6(work$postcode)
  work$address_key    <- address_key(work$postcode, work$huisnummer, work$huisletter, work$toevoeging)

  say("BAG ID check: %s", paste(sprintf("%s = %d", names(table(work$bag_id_status)), as.integer(table(work$bag_id_status))), collapse = "; "))
  if (any(!work$postcode_valid)) say("Invalid postcodes: %d row(s) (they get PC4/national fall-backs only).", sum(!work$postcode_valid))
  work
}

# ---- feature assembly ---------------------------------------------------------

# Attach BAG building characteristics. Matching order:
#   1. nummeraanduiding ID  2. verblijfsobject ID  3. postcode + house number
join_bag <- function(work, bag) {
  bag_cols <- intersect(c("oppervlakte_m2", "bouwjaar", "verblijfsobject_id", "gebruiksdoel"), names(bag))
  out <- work %>% mutate(bag_match = "none", oppervlakte_m2 = NA_real_, bouwjaar = NA_integer_,
                         bag_verblijfsobject_id = NA_character_)
  take <- function(out, key_work, key_bag, label) {
    todo <- which(out$bag_match == "none" & !is.na(out[[key_work]]))
    if (length(todo) == 0) return(out)
    ref <- bag[!is.na(bag[[key_bag]]), ] %>% distinct(.data[[key_bag]], .keep_all = TRUE)
    m <- match(out[[key_work]][todo], ref[[key_bag]])
    hit <- !is.na(m)
    if (any(hit)) {
      rows <- todo[hit]
      out$oppervlakte_m2[rows] <- ref$oppervlakte_m2[m[hit]]
      out$bouwjaar[rows]       <- ref$bouwjaar[m[hit]]
      out$bag_verblijfsobject_id[rows] <- ref$verblijfsobject_id[m[hit]]
      out$bag_match[rows] <- label
    }
    out
  }
  out <- out %>% mutate(.key_na = ifelse(bag_id_kind == "nummeraanduiding", bag_id, NA_character_),
                        .key_vo = ifelse(bag_id_kind == "verblijfsobject", bag_id, NA_character_))
  out <- take(out, ".key_na", "bag_id", "nummeraanduiding")
  out <- take(out, ".key_vo", "verblijfsobject_id", "verblijfsobject")
  if ("address_key" %in% names(bag)) out <- take(out, "address_key", "address_key", "address")
  out %>% select(-.key_na, -.key_vo)
}

# Attach CBS postcode statistics with PC6 -> PC5 -> PC4 fall-back per variable.
join_cbs <- function(work, cbs) {
  vars <- c("female_share", "avg_age", "huishoudgrootte", "rental_share", "income_z", "urban", "woz_eur")
  x <- work %>% mutate(.pc5 = substr(postcode, 1, 5), .pc4 = substr(postcode, 1, 4))
  for (lv in c("pc6", "pc5", "pc4")) {
    t <- cbs[[lv]]
    t <- t[, intersect(c("postcode", vars), names(t))]
    names(t)[-1] <- paste0(names(t)[-1], "__", lv)
    by <- c(pc6 = "postcode", pc5 = ".pc5", pc4 = ".pc4")[[lv]]
    x <- left_join(x, t, by = setNames("postcode", by))
  }
  fallback_notes <- rep("", nrow(x))
  for (v in vars) {
    cols <- paste0(v, "__", c("pc6", "pc5", "pc4"))
    cols <- cols[cols %in% names(x)]
    if (length(cols) == 0) next
    vals <- as.data.frame(x[cols])
    lvl  <- apply(vals, 1, function(r) { i <- which(!is.na(r)); if (length(i)) sub(".*__", "", cols[i[1]]) else NA })
    x[[v]] <- do.call(coalesce, unname(as.list(vals)))
    finest <- sub(".*__", "", cols[1])
    used_coarser <- !is.na(lvl) & lvl != finest
    fallback_notes[used_coarser] <- paste0(fallback_notes[used_coarser], v, "<-", lvl[used_coarser], "; ")
    x[[paste0(v, "__level")]] <- lvl
  }
  x$cbs_fallback <- sub("; $", "", fallback_notes)
  x %>% select(-all_of(grep("__pc[456]$", names(x), value = TRUE)), -.pc5, -.pc4)
}

# Everything together: BAG + CBS -> the eight model features, with imputation
# flags for rows that had no data at any level (so prediction never crashes and
# nothing is filled silently).
assemble_features <- function(work, open_data, bundle) {
  x <- work %>% join_bag(open_data$bag) %>% join_cbs(open_data$cbs)

  yb_levels <- bundle$year_built_levels
  yb_breaks <- bundle$year_built_breaks
  x$year_built <- ifelse(is.na(x$bouwjaar), NA_character_, yb_levels[findInterval(x$bouwjaar, yb_breaks) + 1])

  x <- x %>% mutate(female = female_share, age = avg_age, members = huishoudgrootte,
                    rental = rental_share, nettohh_z = income_z, square_meters = oppervlakte_m2)

  # Fall-backs: population medians from the open data of the year used; the
  # training medians in the bundle are the last resort.
  dflt <- list(
    female        = as.numeric(meta_value(open_data, "default_female_share")),
    age           = as.numeric(meta_value(open_data, "default_avg_age")),
    members       = as.numeric(meta_value(open_data, "default_huishoudgrootte")),
    rental        = as.numeric(meta_value(open_data, "default_rental_share")),
    urban         = as.numeric(meta_value(open_data, "default_urban")),
    nettohh_z     = 0,
    square_meters = as.numeric(meta_value(open_data, "default_square_meters")),
    year_built    = yb_levels[as.integer(meta_value(open_data, "default_year_built_index"))])
  for (f in names(dflt)) if (is.na(dflt[[f]]) || length(dflt[[f]]) == 0) dflt[[f]] <- bundle$training$feature_defaults[[f]]

  imputed <- rep("", nrow(x))
  for (f in bundle$features) {
    miss <- is.na(x[[f]])
    if (any(miss)) {
      x[[f]][miss] <- dflt[[f]]
      imputed[miss] <- paste0(imputed[miss], f, "; ")
    }
  }
  x$imputed_features <- sub("; $", "", imputed)

  say("\nFeature assembly: %d addresses", nrow(x))
  say("  BAG match: %s", paste(sprintf("%s = %d", names(table(x$bag_match)), as.integer(table(x$bag_match))), collapse = "; "))
  say("  CBS: %d with PC6 demographics, %d needed a coarser level, %d with postcode income",
      sum(x$female_share__level == "pc6", na.rm = TRUE), sum(x$cbs_fallback != ""), sum(!is.na(x$income_z)))
  say("  Imputed with population fall-backs: %d row(s)%s", sum(x$imputed_features != ""),
      if (any(x$imputed_features != "")) paste0(" (", paste(names(sort(table(unlist(strsplit(x$imputed_features[x$imputed_features != ""], "; "))), decreasing = TRUE)), collapse = ", "), ")") else "")
  x
}

# ---- model bundle -------------------------------------------------------------

read_model_bundle <- function(path) {
  if (!file.exists(path)) stop("Model file not found: ", path, "\nIt is models/persona_rf.rds in the repository; check repo_dir / model_path in the config chunk.", call. = FALSE)
  b <- readRDS(path)
  stopifnot(is.list(b), inherits(b$forest, "ranger"))
  if (!requireNamespace("ranger", quietly = TRUE)) stop("Package 'ranger' is needed: install.packages('ranger')", call. = FALSE)
  rv <- as.character(packageVersion("ranger"))
  if (!identical(rv, b$training$ranger_version)) {
    say("note: model trained with ranger %s, you have %s. Predictions are normally unaffected.", b$training$ranger_version, rv)
  }
  b
}

# Score assembled features. Returns the four-class prediction (the model's
# target) and the deployed persona after the Erratic -> Comfort fold.
predict_personas <- function(bundle, features) {
  X <- as.data.frame(features[bundle$features])
  X$year_built <- factor(as.character(X$year_built), levels = bundle$year_built_levels)
  for (v in setdiff(bundle$features, "year_built")) X[[v]] <- as.numeric(X[[v]])
  na_cols <- names(X)[colSums(is.na(X)) > 0]
  if (length(na_cols) > 0) stop("Missing values in model features: ", paste(na_cols, collapse = ", "),
                                ". Run assemble_features() first (it fills them and records what it filled).", call. = FALSE)
  p <- predict(bundle$forest, data = X)$predictions
  class4 <- as.character(p)
  folded <- ifelse(class4 %in% names(bundle$fold_rule), unname(bundle$fold_rule[class4]), class4)
  tibble(persona_class4 = class4,
         persona_class4_label = unname(bundle$class_labels[class4]),
         persona = folded,
         persona_label = unname(bundle$class_labels[folded]))
}

print_model_card <- function(bundle) {
  t <- bundle$training
  say("\n==== Persona random forest (model card) ====")
  say("trained %s | n = %d | %d trees | ranger %s", t$date, t$n_obs, t$num_trees, t$ranger_version)
  say("tuning (%s, CV): %s", t$cv_metric, paste(sprintf("%s = %s", names(t$tuning), vapply(t$tuning, as.character, "")), collapse = ", "))
  say("class weights: %s", paste(sprintf("%s = %.2f", names(t$class_weights), t$class_weights), collapse = ", "))
  say("OOB error (four-class, before the Erratic -> Comfort fold): %.3f", t$oob_error)
  say("OOB confusion (rows = true, cols = predicted):")
  print(t$oob_confusion)
  cat("\nper class: ")
  print(round(rbind(recall = t$oob_recall, precision = t$oob_precision, lift = t$oob_lift), 3))
  cat("\nOOB confusion after folding predicted Erratic into Comfort (deployment view):\n")
  print(t$oob_confusion_folded)
  cat("\npermutation importance:\n")
  print(round(t$importance, 4))
  say("deployment rule: %s", paste(sprintf("%s -> %s", names(bundle$fold_rule), bundle$fold_rule), collapse = ", "))
  say("features: %s", paste(bundle$features, collapse = ", "))
  say("=============================================\n")
}
```

# Step 1: Open data (built once, then read from the cache)

```{r open-data}
open_data <- get_open_data(data_dir = path.expand(data_dir), cbs_year = cbs_year, income_year = income_year,
                           cache_dir = path.expand(cache_dir), bag_gpkg = path.expand(bag_gpkg),
                           rebuild = rebuild_cache)
print_open_data_summary(open_data)
```

# Step 2: The trained model

```{r model}
bundle <- read_model_bundle(model_path)
print_model_card(bundle)
```

# Step 3: Your addresses

```{r addresses}
addresses_original <- if (is.data.frame(addresses)) {
  as_tibble(addresses) %>% mutate(across(everything(), as.character))
} else read_address_file(from_repo(addresses), addresses_sheet, addresses_skip)

cat(nrow(addresses_original), "addresses; columns:", paste(names(addresses_original), collapse = " | "), "\n")

# Working copy with clean keys; the original is written back untouched in Step 5.
work <- standardise_addresses(addresses_original, address_col_map)
```

```{r assemble}
features <- assemble_features(work, open_data, bundle)

# Addresses that needed any fall-back, for the record
features %>%
  filter(bag_match != "nummeraanduiding" | cbs_fallback != "" | imputed_features != "") %>%
  count(bag_match, cbs_fallback, imputed_features, name = "addresses") %>%
  arrange(-addresses)
```

# Step 4: Predict

```{r predict}
pred <- predict_personas(bundle, features)
scored <- bind_cols(features, pred)

table(scored$persona_class4_label)     # the forest's four classes
table(scored$persona_label)            # after Erratic -> Comfort

ggplot(scored, aes(x = persona_label)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Predicted persona distribution", x = "Persona", y = "Addresses") +
  theme_minimal()
```

# Step 5: Write the output

Your columns exactly as they were, followed by the model's columns. If one of
the appended names already exists in your file, the appended copy gets `_model`.

```{r write-output}
extra <- scored %>%
  select(bag_id_clean = bag_id, bag_id_status, bag_match, oppervlakte_m2, bouwjaar, year_built,
         female, age, members, rental, urban, nettohh_z, square_meters,
         cbs_fallback, imputed_features,
         persona_class4, persona_class4_label, persona, persona_label)
clash <- names(extra) %in% names(addresses_original)
names(extra)[clash] <- paste0(names(extra)[clash], "_model")
out <- bind_cols(addresses_original, extra)
stopifnot(nrow(out) == nrow(addresses_original))

path_output_csv <- from_repo(path_output_csv)
dir.create(dirname(path_output_csv), showWarnings = FALSE, recursive = TRUE)
write_csv(out, path_output_csv, na = "")
cat("Wrote", path_output_csv, "\n")

if (!is.null(path_output_xlsx)) {
  path_output_xlsx <- from_repo(path_output_xlsx)
  dir.create(dirname(path_output_xlsx), showWarnings = FALSE, recursive = TRUE)
  write_xlsx(list("Personas" = out), path_output_xlsx)
  cat("Wrote", path_output_xlsx, "\n")
}

cat("\nPersonas (after Erratic -> Comfort):\n")
print(table(out$persona_label, useNA = "ifany"))
```
